Direct Entropy Measurement in a Mesoscopic Quantum System

### 2018-01-26 paper/draft.tex
edits...

%%%%%%%%%%  current edits end here  %%%%%%%%%%%%%

@@

To make a  measurement of the quantum dot entropy, the right dot was coupled to the reservoir on the source (left) side, with the drain (right) side of the dot closed. The charge sensing QPC  was tuned to a conductance of ${\sim}e^2/h$. A small DC voltage bias was applied and the measured current was used to calculate the sensor conductance.  Assuming the source barrier is in the limit where single electron transitions ($\Gamma_{In/Out}$) are only broadened by the temperature of the reservoir (discussed further in Fig. \ref{fig:fig2}c, the charge sensor conductance takes the form,
%
\begin{align}
\label{eqn:g}
        g_{sens}(V,\Theta) &= G_0 \tanh\left(\frac{V-V_0(\Theta)}{2\Theta}\right)  \\
                        &\quad + G_1\left[V-V_0(\Theta)\right] + G_2 \nonumber
\end{align}
%
where $\Theta = \frac{k_B T}{\alpha e}$, $\alpha$ is the lever arm for the quantum dot plunger gate, and $G_1$ is determined by the cross capacitance between the sensor gates and the plunger gate.

The temperature of the electron reservoir was varied by applying an AC current through the heating QPC on the left side of the channel. Fluctuations in $g_{sens}$ were measured at $2f_{Heat}$ using a lock-in amplifier; the temperature of the reservoir goes as $dT^2 \sim P^{AC}_{Heat} \sim \sin^2(2\omega t)$. Taking a derivative of Eq. \ref{eqn:g} with respect to temperature gives the change in conductance, $dg_{sens}$, as measured by the lock-in amplifier,
%
\begin{align}
\label{eqn:dg}
        dg_{sens}(V, \Theta) &\propto \left[ \frac{V-V_0(\Theta)}{2\Theta} + \delta \right]\times \\
        				      &\quad\cosh^{-2}\left(\frac{V-V_0(\Theta)}{2\Theta}\right) \nonumber
\end{align}
%
where $\delta=\frac{\partial V_0}{\partial \Theta}$. By using the lever arm to convert $V_0$ to chemical potential $\mu$, and recalling Eq. \ref{eqn:max}, the parameter $\delta$ can be related to the change in entropy,
%
\begin{align}
\label{eqn:delta}
        \delta &= \frac{\partial V_0}{\partial \Theta} = 
        \frac{1}{k_B} \frac{\partial \mu}{\partial T} = 
        -\frac{1}{k_B} \frac{\partial S}{\partial N}
\end{align}
%

>>>>

%%%%%%%%%%%%%%%%%%%%%%%

@@

%%% device setup %%%
Using the gates on the right hand side of Fig~\ref{fig:fig1}a, the quantum dot was tuned such that there is a weak tunnel coupling to the reservoir on the source side with the drain side of the dot closed. The charge sensing QPC was tuned to a conductance of ${\sim}e^2/h$. A small DC voltage bias was applied across the QPC and the measured current was used to calculate the sensor conductance. The reservoir temperature was controlled either by varying the fridge temperature or by Joule heating through the QPC on the left side of the channel. In the latter case, the QPC was fixed at \SI{20}{\kilo\ohms} and biased using an AC or DC current source. The Joule heating strategy gives fast control of the temperature (up to \SI{100}{\hertz}) while keeping the voltage drop across the reservoir negligably small.

%%% figs 2a and b %%%
Assuming the source barrier is in the limit where single electron transitions ($\Gamma_{In/Out}$) are only broadened by the temperature of the reservoir (discussed further in Fig. \ref{fig:fig2}c, the charge sensor conductance takes the form,
%
\begin{align}
\label{eqn:g}
        g_{sens}(V,\Theta) &= G_0 \tanh\left(\frac{V-V_0(\Theta)}{2\Theta}\right)  \\
                        &\quad + G_1\left[V-V_0(\Theta)\right] + G_2 \nonumber
\end{align}
%
where $\Theta = \frac{k_B T}{\alpha e}$, $\alpha$ is the lever arm for the quantum dot plunger gate, and $G_1$ is determined by the cross capacitance between the sensor gates and the plunger gate.

%%% figs 2c and d%%%
Fluctuations in $g_{sens}$ were measured at $2f_{Heat}$ using a lock-in amplifier; the temperature of the reservoir goes as $dT^2 \sim P^{AC}_{Heat} \sim \sin^2(2\omega t)$. Taking a derivative of Eq. \ref{eqn:g} with respect to temperature gives the change in conductance, $dg_{sens}$, as measured by the lock-in amplifier,
%
\begin{align}
\label{eqn:dg}
        dg_{sens}(V, \Theta) &\propto \left[ \frac{V-V_0(\Theta)}{2\Theta} + \delta \right]\times \\
        				      &\quad\cosh^{-2}\left(\frac{V-V_0(\Theta)}{2\Theta}\right) \nonumber
\end{align}
%
where $\delta=\frac{\partial V_0}{\partial \Theta}$. By using the lever arm to convert $V_0$ to chemical potential $\mu$, and recalling Eq. \ref{eqn:max}, the parameter $\delta$ can be related to the change in entropy,
%
\begin{align}
\label{eqn:delta}
        \delta &= \frac{\partial V_0}{\partial \Theta} = 
        \frac{1}{k_B} \frac{\partial \mu}{\partial T} = 
        -\frac{1}{k_B} \frac{\partial S}{\partial N}
\end{align}
%
